---
title: "Week 02: COVID Visualization Activity"
date: "September, 11 2023"
author: "**`[[`**Brianna Li**`]]`**"
format: gfm
---

Today, we'll be working with a data set related to COVID.
This data is based on data from the the [COVID Tracking Project](https://covidtracking.com/).
I cleaned up this data and also added total populations from the 2020 for each of the relevant categories.
Note, due to differences in the way race and ethnicity are encoded in the census as compared to the the COVID Tracking Project, the population counts for LatinX may be somewhat inaccurate.

```{r load_data}
library(tidyverse)
load("CRDT Data.RData")
ls()
```

I've include 4 different data sets.
They all have the same data but have it represented in different ways.
Try using the different data sets and see which ones are good for making which plots.


```{r first_plot}
ggplot(covid_data_count, aes(x = date, y = Cases)) + geom_point()
```



If you want to only look at a specific state, you can do it like this.
For now, see what you can do just using `ggplot`.

```{r}
covid_data_count |> 
  filter(state == "MA") |> 
  ggplot(aes(x = date, y = Cases, color = race)) + geom_line()
```

1. After making and refining some plots, pick one and save the code and describe what you observe?

```{r}
# add death rate
covid_data_simple <- covid_data_simple |>
  mutate(death_rate = (Deaths / population) * 100)  # Ensure population is the total for each race group

# plot covid-19 death rate by race
covid_data_simple |>
  ggplot(aes(x = date, y = death_rate, color = race)) + 
  geom_line() +
  labs(title = "COVID-19 Death Rate by Race", 
       x = "Date", 
       y = "Death Percentage (%)",
       color = "Race")
```
```{r}
# add death rate
covid_data_simple <- covid_data_simple |>
  mutate(hosp_rate = (Hosp / population) * 100)  # Ensure population is the total for each race group

# plot covid-19 death rate by race
covid_data_simple |>
  ggplot(aes(x = date, y = hosp_rate, color = race)) + 
  geom_line() +
  labs(title = "COVID-19 Hospitalization Rate by Race", 
       x = "Date", 
       y = "Hospitalization Percentage (%)",
       color = "Race")
```
```{r}

# add death rate and hospitalization rate to covid data simple
covid_data_simple <- covid_data_simple |>
  mutate(
    death_rate = (Deaths / population) * 100,  # Death percentage
    hosp_rate = (Hosp / population) * 100     # Hospitalization percentage
  )

# reshape data to long format
covid_data_long <- covid_data_simple |>
  pivot_longer(cols = c(death_rate, hosp_rate), 
               names_to = "rate_type", 
               values_to = "rate_value")

# plot covid-19 death and hospitalization rates by race
ggplot(covid_data_long, aes(x = date, y = rate_value, color = race, linetype = rate_type)) + 
  geom_line() +
  scale_linetype_manual(values = c("death_rate" = "dotted", "hosp_rate" = "solid"))
  labs(
    title = "COVID-19 Death and Hospitalization Rates by Race", 
    x = "Date", 
    y = "Rate Percentage (%)",
    color = "Race",
    linetype = "Rate Type"
  )

```

2. What conclusions can you draw?
```{md}
Before October, the death rate was higher that the hospitalization rate for all races. In addition, the rate of hospitalization for Blacks and Latinx were the highest and also the most similar. The death rate was highest for the black population. In October, there was a jump in hospitalization rate. This was also around the time that the COVID-19 vaccination was approved.
```

3. What were you not able to do due to not having the coding knowledge?
```{md}
I wanted to create a graph that would show the funnel/rate of positive tests to hospitalization to death to see the conversion rates.
```
4. What made previous plots you made hard to interpret/understand? What did you do to improve/change them to ease interpretation?
5. What other data would be useful to better understand this data?